---
title: "treewell_publication"
author: "Erika Lee"
date: "2025-05-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lterdatasampler)
library(dplyr)
library(ggplot2)
library(readr)
library(readxl)
library(lubridate)
library(plotly)
library(openxlsx)
library(plotly)
library(rstatix)
library(htmlwidgets)
library(RColorBrewer)
library(patchwork)
library(ggpubr)
library(suncalc)

setwd("/Volumes/wcnr-network/Research/Kampf/Private/field_data")
```

# Raw Data Processing

## Tree Temperature Data

PSZ Burned

```{r}
#read in temperature data
pers_b_tree_composite=read_excel("nsf/105E_pers_burned/105E_pers_b_composite.xlsx")

#filter temp data to full data frame
pers_b_tree_fulldata <- pers_b_tree_composite %>%
  mutate(datetime = ymd_hms(TIMESTAMP)) %>%
  select(datetime, everything()) %>%
  select(-c(TIMESTAMP, RECORD))

#adding in an aspect column - WITH ONLY 14 SENSORS!
pers_b_aspects <- pers_b_tree_fulldata %>%
  pivot_longer(cols = starts_with("Temp"),
               names_to = "sensor_id",
               values_to = "temp") %>%
  #add in aspect column
  mutate(aspect = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(1)",
      "Temp_C_Avg(5)",
      "Temp_C_Avg(9)",
      "Temp_C_Avg(13)", "Temp_C_Avg(15)"
    ),
    "north",
    "other"
  )) %>%
  mutate(aspect = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(2)",
      "Temp_C_Avg(6)",
      "Temp_C_Avg(10)",
      "Temp_C_Avg(14)", "Temp_C_Avg(16)"),
    "south",
    aspect
  )) %>%
  mutate(aspect = ifelse(
    sensor_id %in% c("Temp_C_Avg(3)", "Temp_C_Avg(7)", "Temp_C_Avg(11)"),
    "charred",
    aspect
  )) %>%
  mutate(aspect = ifelse(
    sensor_id %in% c("Temp_C_Avg(4)", "Temp_C_Avg(8)", "Temp_C_Avg(12)"),
    "uncharred",
    aspect
  )) %>%
  #adding in sensor name column
  mutate(tree_name = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(1)",
      "Temp_C_Avg(2)",
      "Temp_C_Avg(3)",
      "Temp_C_Avg(4)"
    ),
    "gb_w",
    "other"
  )) %>%
  mutate(tree_name = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(5)",
      "Temp_C_Avg(6)",
      "Temp_C_Avg(7)",
      "Temp_C_Avg(8)"
    ),
    "gb_e",
    tree_name
  )) %>%
  mutate(tree_name = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(9)",
      "Temp_C_Avg(10)",
      "Temp_C_Avg(11)",
      "Temp_C_Avg(12)"
    ),
    "gb_s",
    tree_name
  )) %>%
  mutate(tree_name = ifelse(
    sensor_id %in% c("Temp_C_Avg(13)", "Temp_C_Avg(14)"),
    "db_n",
    tree_name
  )) %>%
  mutate(tree_name = ifelse(
    sensor_id %in% c("Temp_C_Avg(15)", "Temp_C_Avg(16)"),
    "db_s",
    tree_name
  ))
```

PSZ Unburned

```{r}
#read in temperature data
pers_ub_tree_composite=read_excel("nsf/105E_pers_unburned/105E_pers_ub_composite.xlsx")

#filter temp data to full data frame
pers_ub_tree_fulldata <- pers_ub_tree_composite %>%
  mutate(datetime = as.POSIXct(TIMESTAMP, format=("%Y-%m-%d %H:%M:%S")), datetime = force_tz(datetime, tz = 'MST')) %>%
  select(datetime, everything()) %>%
  select(-c(TIMESTAMP, RECORD))

#adding in an aspect column - WITH ONLY 14 SENSORS!
pers_ub_aspects <- pers_ub_tree_fulldata %>%
  pivot_longer(cols = starts_with("Temp"),
               names_to = "sensor_id",
               values_to = "temp") %>%
  #add in aspect column
  mutate(aspect = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(1)",
      "Temp_C_Avg(3)"),
    "north",
    "other"
  )) %>%
  mutate(aspect = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(2)",
      "Temp_C_Avg(4)"),
    "south",
    aspect
  )) %>%
  #adding in sensor name column
  mutate(tree_name = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(1)",
      "Temp_C_Avg(2)"
    ),
    "ub_n",
    "other"
  )) %>%
  mutate(tree_name = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(3)", 
      "Temp_C_Avg(4)"
    ),
    "ub_s",
    tree_name
  )) 
```

TSZ Burned

```{r}
#read in temperature data
trans_b_tree_composite=read_excel("nsf/105E_trans_burned/105E_trans_b_composite.xlsx")

#filter temp data to full data frame
trans_b_tree_fulldata <- trans_b_tree_composite %>%
  mutate(datetime = ymd_hms(TIMESTAMP)) %>%
  select(datetime, everything()) %>%
  select(-c(TIMESTAMP, Record)) %>%
  na.omit()

#adding in an aspect column - 
trans_b_aspects <- trans_b_tree_fulldata %>%
  pivot_longer(cols = starts_with("Temp"),
               names_to = "sensor_id",
               values_to = "temp") %>%
  #add in aspect column
  mutate(aspect = ifelse(
    sensor_id %in% c("Temp_C_Avg(1)", "Temp_C_Avg(5)", "Temp_C_Avg(9)", "Temp_C_Avg(11)"),
    "north",
    "other"
  )) %>%
mutate(aspect = ifelse(
    sensor_id %in% c("Temp_C_Avg(2)", "Temp_C_Avg(6)", "Temp_C_Avg(10)", "Temp_C_Avg(12)"),
    "south",
    aspect)) %>%
  mutate(aspect = ifelse(
    sensor_id %in% c("Temp_C_Avg(3)", "Temp_C_Avg(7)"),
    "charred",
    aspect)) %>%
    mutate(aspect = ifelse(
    sensor_id %in% c("Temp_C_Avg(4)", "Temp_C_Avg(8)"),
    "uncharred",
    aspect)) %>%
  #adding in sensor name column
  mutate(tree_name = ifelse(
    sensor_id %in% c("Temp_C_Avg(1)", "Temp_C_Avg(2)", "Temp_C_Avg(3)", "Temp_C_Avg(4)"),
    "gb_n",
    "other"
  )) %>%
mutate(tree_name = ifelse(
    sensor_id %in% c("Temp_C_Avg(5)", "Temp_C_Avg(6)", "Temp_C_Avg(7)", "Temp_C_Avg(8)"),
    "gb_s",
    tree_name)) %>%
  mutate(tree_name = ifelse(
    sensor_id %in% c("Temp_C_Avg(9)", "Temp_C_Avg(10)"),
    "db_n",
    tree_name)) %>%
    mutate(tree_name = ifelse(
    sensor_id %in% c("Temp_C_Avg(11)", "Temp_C_Avg(12)"),
    "db_s",
    tree_name))
```

TSZ Unburned

```{r}
trans_ub_tree_composite=read_excel("nsf/105E_trans_unburned/105E_trans_ub_composite.xlsx")

#filter temp data to full data frame
trans_ub_tree_fulldata <- trans_ub_tree_composite %>%
  mutate(datetime = as.POSIXct(TIMESTAMP, format = ("%Y-%m-%d %H:%M:%S")),
         datetime = force_tz(datetime, tz = 'MST'))%>%
  select(datetime, everything()) %>%
  select(-c(TIMESTAMP,RECORD))

#adding in an aspect column - WITH ONLY 14 SENSORS!
trans_ub_aspects <- trans_ub_tree_fulldata %>%
  pivot_longer(cols = starts_with("Temp"),
               names_to = "sensor_id",
               values_to = "temp") %>%
  #add in aspect column
  mutate(aspect = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(1)",
      "Temp_C_Avg(3)"),
    "north",
    "other"
  )) %>%
  mutate(aspect = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(2)",
      "Temp_C_Avg(4)"),
    "south",
    aspect
  )) %>%
  #adding in sensor name column
  mutate(tree_name = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(1)",
      "Temp_C_Avg(2)"
    ),
    "ub_w",
    "other"
  )) %>%
  mutate(tree_name = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(3)", 
      "Temp_C_Avg(4)"
    ),
    "ub_e",
    tree_name
  )) 
```

## Weather Station Data

```{r}
transb_wx_15min_r <- read_csv('nsf/trans_burned/trans_burned_wx_15min_r.csv') %>%
  #have to force the MST otherwise it recognizes it as UTC
  mutate(datetime = force_tz(datetime, tz = 'MST'))

transub_wx_15min_r <- read_csv('nsf/trans_unburned/trans_unburned_wx_15min_r.csv') %>%
  mutate(datetime = force_tz(datetime, tz = 'MST'))

persb_wx_hourly_r <- read_csv('nsf/pers_burned/pers_burned_wx_hourly_r.csv') %>%
  mutate(datetime = force_tz(datetime, tz = 'MST'))

persub_wx_15min_r <- read_csv('nsf/pers_unburned/pers_unburned_wx_15min_r.csv') %>%
  mutate(datetime = force_tz(datetime, tz = 'MST'))
```

# Data Wrangling

Setting a constant time frame for data for period of observation

Tree temperature data

```{r}
filtered_pers_b_aspects <- pers_b_aspects %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST")) %>%
  filter(datetime >= as.POSIXct("2024-02-02 00:00:00", tz = "MST") & 
         datetime < as.POSIXct("2024-07-01 00:00:00", tz = "MST"))

filtered_pers_ub_aspects <- pers_ub_aspects %>%
 mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST")) %>%
  filter(datetime >= as.POSIXct("2024-02-02 00:00:00", tz = "MST") & 
         datetime < as.POSIXct("2024-07-01 00:00:00", tz = "MST"))

filtered_trans_b_aspects <- trans_b_aspects %>%
 mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST")) %>%
  filter(datetime >= as.POSIXct("2024-02-02 00:00:00", tz = "MST") & 
         datetime < as.POSIXct("2024-07-01 00:00:00", tz = "MST"))

filtered_trans_ub_aspects <- trans_ub_aspects %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST")) %>%
  filter(datetime >= as.POSIXct("2024-02-02 00:00:00", tz = "MST") & 
         datetime < as.POSIXct("2024-07-01 00:00:00", tz = "MST"))
```

Weather station data

```{r}
#setting consistent timeframe for weather data
#filtered weather dataframes
filtered_persb_wx_hourly <- persb_wx_hourly_r %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST")) %>%
  filter(datetime >= as.POSIXct("2024-02-02 00:00:00", tz = "MST") & 
         datetime < as.POSIXct("2024-07-01 00:00:00", tz = "MST"))

filtered_persub_wx_15min <- persub_wx_15min_r %>%
  select(datetime, everything()) %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST")) %>%
  filter(datetime >= as.POSIXct("2024-02-02 00:00:00", tz = "MST") & 
         datetime < as.POSIXct("2024-07-01 00:00:00", tz = "MST"))
  
filtered_transb_wx_15min <- transb_wx_15min_r %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST")) %>%
  filter(datetime >= as.POSIXct("2024-02-02 00:00:00", tz = "MST") & 
         datetime < as.POSIXct("2024-07-01 00:00:00", tz = "MST"))

filtered_transub_wx_15min <- transub_wx_15min_r %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST")) %>%
  filter(datetime >= as.POSIXct("2024-02-02 00:00:00", tz = "MST") & 
         datetime < as.POSIXct("2024-07-01 00:00:00", tz = "MST"))
```

Adding in Solar Exposure Data

```{r}
# Specify PSZ and TSZ locations
##pers_b - from 105E location!
p_b_lat <- 40.56379284
p_b_lon <- -105.867215
##pers_ub - from 105E locations
p_ub_lat <- 40.5638075
p_ub_lon <- -105.8705715

##trans_b - from 105E locations!
t_b_lat <- 40.60600629
t_b_lon <- -105.8333335

##trans_ub - from 105E locations!
t_ub_lat <- 40.568396
t_ub_lon <- -105.587383

## use the p_b_lat/long since all of these dates/times are almost identical for sunrise/sunset

# Create a date sequence from February 2nd to May 24th, 2024 for solar exposure
solex_dates <- seq(as.Date("2024-02-02"), as.Date("2024-07-01"), by = "day")

# Get solar times for each day
p_b_solar_times <- getSunlightTimes(date = solex_dates, lat = p_b_lat, lon = p_b_lon, tz = "MST") %>%
  mutate(
    day = as.Date(date),  # Ensure 'day' is of Date type for joining
    sunrise = with_tz(as.POSIXct(sunrise), tzone = "MST"),  # Set MST timezone
    sunset = with_tz(as.POSIXct(sunset), tzone = "MST")     # Set MST timezone
  )
```

Adding in Burn Condition

```{r}
#adding in a burn status column to the aspect dataframes
filtered_pers_b_aspects <- filtered_pers_b_aspects %>%
  mutate(burn_status = case_when(
    startsWith(tree_name, "gb") ~ "green burn",
    startsWith(tree_name, "db") ~ "dead burn",
    TRUE ~ NA_character_  # Optional: handle other cases if needed
  )) %>%
  #adding a zone column
  mutate(zone = "PSZ") %>%
  mutate(day = as.POSIXct(format(datetime, "%Y-%m-%d")),
         month = as.numeric(format(datetime, "%m"))) %>%
  mutate(snow_phase = case_when(
      day >= as.Date("2024-05-06") ~ "ablation",
      TRUE ~ "accumulation"
    )) %>%
  select(datetime, day, month, zone, snow_phase, burn_status, sensor_id, tree_name, everything()) %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST"))

filtered_pers_ub_aspects <- filtered_pers_ub_aspects %>%
  mutate(burn_status = case_when(
    startsWith(tree_name, "ub") ~ "live",
    TRUE ~ NA_character_  # Optional: handle other cases if needed
  )) %>%
  #adding a zone column
  mutate(zone = "PSZ") %>%
  mutate(day = as.POSIXct(format(datetime, "%Y-%m-%d")),
         month = as.numeric(format(datetime, "%m"))) %>%
  mutate(snow_phase = case_when(
      day >= as.Date("2024-05-06") ~ "ablation",
      TRUE ~ "accumulation"
    )) %>%
  select(datetime, day, month, zone, snow_phase, burn_status, sensor_id, tree_name, everything()) %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST"))

filtered_trans_b_aspects <- filtered_trans_b_aspects %>%
  mutate(burn_status = case_when(
    startsWith(tree_name, "gb") ~ "green burn",
    startsWith(tree_name, "db") ~ "dead burn",
    TRUE ~ NA_character_  # Optional: handle other cases if needed
  )) %>%
  #adding a zone column
  mutate(zone = "TSZ") %>%
  mutate(day = as.POSIXct(format(datetime, "%Y-%m-%d")),
         month = as.numeric(format(datetime, "%m"))) %>%
  mutate(snow_phase = case_when(
      day >= as.Date("2024-04-01") ~ "ablation",
      TRUE ~ "accumulation"
    )) %>%
  select(datetime, day, month, zone, snow_phase, burn_status, sensor_id, tree_name, everything()) %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST"))

filtered_trans_ub_aspects <- filtered_trans_ub_aspects %>%
  mutate(burn_status = case_when(
    startsWith(tree_name, "ub") ~ "live",
    TRUE ~ NA_character_  # Optional: handle other cases if needed
  )) %>%
  mutate(zone = "TSZ") %>%
  mutate(day = as.POSIXct(format(datetime, "%Y-%m-%d")),
         month = as.numeric(format(datetime, "%m"))) %>%
  mutate(snow_phase = case_when(
      day >= as.Date("2024-04-01") ~ "ablation",
      TRUE ~ "accumulation"
    )) %>%
  select(datetime, day, month, zone, snow_phase, burn_status, sensor_id, tree_name, everything()) %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST"))
```

## NS Dataframe

```{r}
#creating a dataframe for each tree
persb_gb_e_diff_data <- filtered_pers_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'gb_e') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'gb_e')

persb_gb_s_diff_data <- filtered_pers_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'gb_s') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'gb_s')

persb_db_n_diff_data <- filtered_pers_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'db_n') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'db_n')

persb_db_s_diff_data <- filtered_pers_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'db_s') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'db_s')

#creating a combined new dataframe
persb_gbdb_combined_data <- bind_rows(persb_gb_e_diff_data, persb_gb_s_diff_data, persb_db_n_diff_data, persb_db_s_diff_data)
```

```{r}
#creating a dataframe for each tree
transb_gb_n_diff_data <- filtered_trans_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'gb_n') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'gb_n')

transb_gb_s_diff_data <- filtered_trans_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'gb_s') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'gb_s')

transb_db_n_diff_data <- filtered_trans_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'db_n') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'db_n')

transb_db_s_diff_data <- filtered_trans_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'db_s') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'db_s')

#creating a combined new dataframe
transb_gbdb_combined_data <- bind_rows(transb_gb_n_diff_data, transb_gb_s_diff_data, transb_db_n_diff_data, transb_db_s_diff_data)
```

```{r}
#creating a full dataframe for both pers and transitional data for later plotting of the south-north comparison

##individually editing pers and trans combined data before combining both dataframes
persb_gbdb_combined_data <- persb_gbdb_combined_data %>%
  mutate(burn_status = case_when(
    startsWith(treename, "gb") ~ "green_burned",
    startsWith(treename, "db") ~ "dead_burned",
    TRUE ~ NA_character_  # Handle other cases if needed
  )) %>%
  mutate(zone = "persistent")

transb_gbdb_combined_data <- transb_gbdb_combined_data %>%
  mutate(burn_status = case_when(
    startsWith(treename, "gb") ~ "green_burned",
    startsWith(treename, "db") ~ "dead_burned",
    TRUE ~ NA_character_  # Handle other cases if needed
  )) %>%
  mutate(zone = "transitional")

#creating a full combined dataset
full_ns_comparison_data <- bind_rows(persb_gbdb_combined_data, transb_gbdb_combined_data)
```

```{r}
# important step, I use these later to create the full_ns_dataframe
#creating a combined new dataframe from original dataframes with individual tree temps
full_burned_combined_data <- filtered_pers_b_aspects %>%
  full_join(filtered_trans_b_aspects)

full_unburned_combined_data <- filtered_pers_ub_aspects %>%
  full_join(filtered_trans_ub_aspects)
```

Clean Dataframes

```{r}
#persistent burned 
full_pers_ns_data <- full_burned_combined_data %>%
  filter(!(aspect %in% c("charred", "uncharred"))) %>%
  filter(zone == "PSZ") %>%
  #remove gb_w tree, so there is two gb and two db
  filter(!(tree_name == "gb_w")) %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, snow_phase, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop')

final_persb_ns_temps <- full_pers_ns_data %>%
  group_by(datetime, day, month, burn_status, zone, snow_phase) %>%
  pivot_wider(names_from = aspect, values_from = mean_temp) %>%
  rename(north_temp = north, south_temp = south) %>%
  summarize(north_temp, south_temp)
```

```{r}
#persistent unburned final dataset
full_pers_ub_ns_data <- full_unburned_combined_data %>%
  mutate(burn_status = "live") %>%
  filter(zone == "PSZ") %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, snow_phase, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop')

final_pers_ub_ns_temps <- full_pers_ub_ns_data %>%
  group_by(datetime, day, month, burn_status, zone, snow_phase) %>%
  pivot_wider(names_from = aspect, values_from = mean_temp) %>%
  rename(north_temp = north, south_temp = south) %>%
  summarize(north_temp, south_temp)
```

```{r}
#creating a full n/s green burned, db, ub dataframe to do comparisons on later
## USE THIS DATAFRAME FOR FURTHER DATA ANALYSIS
full_pers_ns_data <- final_persb_ns_temps %>%
  full_join(final_pers_ub_ns_temps)
```

```{r}
#transitional burned 
full_trans_ns_data <- full_burned_combined_data %>%
  filter(!(aspect %in% c("charred", "uncharred"))) %>%
  filter(zone == "TSZ") %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, snow_phase, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop')

final_transb_ns_temps <- full_trans_ns_data %>%
  group_by(datetime, day, month, burn_status, zone, snow_phase) %>%
  pivot_wider(names_from = aspect, values_from = mean_temp) %>%
  rename(north_temp = north, south_temp = south) %>%
  summarize(north_temp, south_temp)
```

```{r}
#tranitional unburned final dataset
full_trans_ub_ns_data <- full_unburned_combined_data %>%
  mutate(burn_status = "live") %>%
  filter(zone == "TSZ") %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, snow_phase, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop')

final_trans_ub_ns_temps <- full_trans_ub_ns_data %>%
  group_by(datetime, day, month, burn_status, zone, snow_phase) %>%
  pivot_wider(names_from = aspect, values_from = mean_temp) %>%
  rename(north_temp = north, south_temp = south) %>%
  summarize(north_temp, south_temp)
```

```{r}
#creating a full n/s green burned, db, ub dataframe to do comparisons on later
## USE THIS DATAFRAME FOR FURTHER DATA ANALYSIS
full_trans_ns_data <- final_transb_ns_temps %>%
  full_join(final_trans_ub_ns_temps)
```

## Burn Condition Dataframe

```{r}
#persistent burned 
full_pers_gbdb_data <- full_burned_combined_data %>%
  filter(!(aspect %in% c("charred", "uncharred"))) %>%
  filter(zone == "PSZ") %>%
  #remove gb_w tree
  filter(!(tree_name == "gb_w")) %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, snow_phase, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop') %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST"))

final_pers_gbdb_temps <- full_pers_gbdb_data %>%
  group_by(datetime, day, month, aspect, zone, snow_phase) %>%
  pivot_wider(names_from = burn_status, values_from = mean_temp) %>%
  rename(gb_temp = "green burn", db_temp = "dead burn") %>%
  summarize(gb_temp, db_temp) %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST"))
```

```{r}
#persistent unburned final dataset
full_pers_ub_data <- full_unburned_combined_data %>%
  mutate(burn_status = "live") %>%
  filter(zone == "PSZ") %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, snow_phase, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop') %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST"))

final_pers_ub_temps <- full_pers_ub_data %>%
  rename(ub_temp = mean_temp) %>%
  select(datetime, day, month, aspect, zone, snow_phase, ub_temp) %>%
  mutate(
    datetime = floor_date(datetime, unit = "15 minutes")  # Rounds down to nearest 15 minutes
  )
```

```{r}
#transitional burned 
full_trans_gbdb_data <- full_burned_combined_data %>%
  filter(!(aspect %in% c("charred", "uncharred"))) %>%
  filter(zone == "TSZ") %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, snow_phase, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop') %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST"))

final_trans_gbdb_temps <- full_trans_gbdb_data %>%
  group_by(datetime, day, month, aspect, zone, snow_phase) %>%
  pivot_wider(names_from = burn_status, values_from = mean_temp) %>%
  rename(gb_temp = "green burn", db_temp = "dead burn") %>%
  summarize(gb_temp, db_temp)
```

```{r}
#transitional unburned final dataset
full_trans_ub_data <- full_unburned_combined_data %>%
  mutate(burn_status = "live") %>%
  filter(zone == "TSZ") %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, snow_phase, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop') %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST"))

final_trans_ub_temps <- full_trans_ub_data %>%
  rename(ub_temp = mean_temp) %>%
  select(-burn_status) %>%
  mutate(
    datetime = floor_date(datetime, unit = "15 minutes")  # Rounds down to nearest 15 minutes
  )
```

Combining burned and unburned dataframes

```{r}
final_pers_gbdb_temps <- final_pers_gbdb_temps %>%
  mutate(datetime = force_tz(datetime, tz = "MST"))

final_pers_ub_temps <- final_pers_ub_temps %>%
  mutate(datetime = force_tz(datetime, tz = "MST"))

full_pers_gbdbub_data <- final_pers_gbdb_temps %>%
  left_join(final_pers_ub_temps)

full_trans_gbdbub_data <- final_trans_gbdb_temps %>%
  left_join(final_trans_ub_temps)
```

```{r}
full_15min_gbdbub <- rbind(full_pers_gbdbub_data, full_trans_gbdbub_data)
```
